---
title: "IPD Meta-Analysis: load and impute"

date:  "`r Sys.Date()`"

output: pdf_document


biometrician: |
  | Svenja Seide
  | seide@imbi.uni-heidelberg.de
    
contacts: |
  | Anand R
  | anand.rubanagarvas@gmail.com
  | 

data-source: |
  | single excel file per study
  
data-supplier: |
  | Anand R
  
data-date: |  
  | "`r Sys.Date()`"

abstract: > 
  To be added
    
draft: true

header-includes:
  - \usepackage{float}
  
---
This Rmd file is heavily edited by me. Not the same copy that Svenja sent
```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = FALSE, 
    cache = TRUE
)
```

```{r dependencies, include=FALSE}
pacman::p_load(haven,lme4,foreign,tidyverse,jomo,mitml,kableExtra,labelled,
               mice,ggmice,lattice,pan,mitools,gridExtra, easystats)

options(print_format="tex")
```

<!-- Use the command: -->
<!-- \blscape -->
<!-- to enter Landscape mode. Use: -->
<!-- \elscape -->
<!-- To exit landscape mode and enter portrait mode again. -->

<!--  # Read the data in --->

```{r Load files}

#load files
files <- (Sys.glob("/Users/personal/Dropbox/Atherosclerosis and iron/IPD meta-analysis/spssfiles/*.sav")) 
data <- lapply(files, function(x) read_sav(x)) #loop function to read files

#read in names of data files
aa <- strsplit(files, "_")  #split based on delimiter
aa <- unlist(lapply(aa, function(x) x[2]))
aa <- unlist(strsplit(aa, "data.sav")) #remove last part of file name

names(data) <- aa #make the names of stored files as row names of the datafile


data_set_ma <- list()
for(i in 1:19){
  data_set_ma[[i]] <- cbind(names(data[i]), data[[i]]$imt, 
                            data[[i]]$iron, data[[i]]$ferritin, 
                            data[[i]]$transferrin, data[[i]]$tfsat, 
                            data[[i]]$age, data[[i]]$gender, data[[i]]$diabetes, 
                            data[[i]]$hypertension, data[[i]]$creatinine, data[[i]]$hdlc, 
                            data[[i]]$ldlc, data[[i]]$triacylglycerols, data[[i]]$bmi, 
                            data[[i]]$smoking, data[[i]]$crp, 
                            data[[i]]$thalassemia,
                            data[[i]]$hemochromatosis, data[[i]]$ckd,data[[i]]$hb,data[[i]]$sbp, data[[i]]$dbp)
}


data_frame <- as.data.frame(do.call(rbind, data_set_ma))
colnames(data_frame) <- c("study", "imt", "iron", "ferritin", "transferrin", 
                          "tfsat", "age", "gender", "diabetes", "hypertension", 
                          "creatinine", "hdlc", "ldlc", "triacylglycerols", 
                          "bmi", "smoking", "crp","thalassemia", 
                          "hemochromatosis", "ckd","hb","sbp","dbp")

```


<!-- # Data mutations --->

```{r}
#data mutations
data_frame <-data_frame |>
   mutate(inflammation=case_when(crp<1~ "0",
                                 crp>=1& crp<3~ "1",
                                 crp>=3~ "2"),
          across(c(imt, iron, ferritin, transferrin, tfsat, age, creatinine, hdlc, ldlc, triacylglycerols, bmi, crp, hb, sbp, dbp),
                as.numeric))|>
  mutate(anemia = case_when(
      age>=15 & gender==0 & hb<13.0 & hb>0 ~ "1",
      age>=15 & gender==1 & hb<12.0 & hb>0~ "1",
      age<5 & hb<11.0 & hb>0 ~ "1",
      age>=5 & age<12 & hb<11.5 & hb>0~ "1",
      age>=12 & age<15 & hb<12.0 & hb>0~ "1",
      age>=15 & gender==0 & hb>=13.0 ~ '0',
      age>=15 & gender==1 & hb>=12.0 ~ '0',
      age<5 & hb>=11.0 ~'0',
      age>=5 & age< 12 & hb>=11.5 ~ '0',
      age>=12 & age<15 & hb>=12.0 ~ '0'),
      age_gp=case_when(age>=18~"Adults",
                          age<18~"Children and Adolescents"),
      age_cat=case_when(age<13~"Children",
                       between(age,13,17.9)~"Adolescents",
                       between(age,18,64.9)~"Adults",
                       age>=65~"Older Adults"),
      hi_ferritin=case_when(tfsat>45 & gender==0 & ferritin>300 |
                                             tfsat>45 & gender==1 & ferritin>200 ~ "1",
                                           tfsat<=45 & gender==0 & ferritin>300 |
                                             tfsat<=45 & gender==1 & ferritin>200 ~ "2"),
      across(c(study, gender, diabetes, hypertension, smoking, thalassemia, hemochromatosis, ckd,
              inflammation, anemia, age_gp,age_cat,hi_ferritin),as.factor)) 

#no filters
unfilt_df <-data_frame
saveRDS(unfilt_df,file=here::here("data_files","unfilt_df.RDS"))
count_df <- unfilt_df%>%
  summarize(
    high_imt = sum(imt > 1.5, na.rm = TRUE),
    high_crp = sum(crp>10, na.rm = TRUE),
    high_ferritin = sum(hemochromatosis=="0"&hi_ferritin == "1", na.rm = TRUE)
  )

saveRDS(count_df,file=here::here("data_files","count_df.RDS"))

#imt filtered
data_frame_sg <- data_frame
data_frame <-data_frame |>
  filter(imt<=1.5|is.na(imt),
         hemochromatosis!="1"|is.na(hemochromatosis),
         thalassemia!="1"|is.na(thalassemia))

data_frame1 <- data_frame|> 
  filter(hi_ferritin!="1"|is.na(hi_ferritin),
         crp<=10|is.na(crp))
  
saveRDS(data_frame,file=here::here("data_files","data_frame.RDS"))
saveRDS(data_frame1,file=here::here("data_files","data_frame1.RDS"))
saveRDS(data_frame_sg,file=here::here("data_files","data_frame_sg.RDS"))
```

\clearpage

# Complete Case Analysis


all analyses are performed without imputation (complete case analysis), this means that all patients with at least one missing covariate that is used in the analysis are excluded. As a consequence, studies that systematically didn't include a variable cannot be used in the analysis.

## iron

The model with the additional adjustment variables cannot be estimated as no observation is included. 

```{r Function-complete cases, results="asis"}
#normalizing variables by scale function
df.comp_adults <- data_frame |>
  filter(age>=18) |>
  dplyr::mutate(iron = scale(iron), 
         ferritin = scale(ferritin), 
         transferrin = scale(transferrin), 
         tfsat = scale(tfsat), 
         age = scale(age), 
         creatinine = scale(creatinine), 
         hdlc = scale(hdlc), 
         ldlc = scale(ldlc), 
         triacylglycerols = scale(triacylglycerols),
         bmi = scale(bmi), 
         crp = scale(crp),
         sbp= scale(sbp),
         dbp=scale(dbp),
         hb=scale(hb))


library(lmerTest)

#define the function to carrout regression and summarize the result
regression= function(model,var) {
  form = reformulate(c(model,var), response = imt)
  lmer(form, data = df.comp_adults) |> 
    model_parameters()
}
#define variables
imt="imt" 
iron="iron"
ferritin="ferritin"
transferrin="transferrin"
tfsat="tfsat"

#define models
model1 = c("1","(1 | study)")
model2=c(model1,"age","gender","diabetes","bmi","smoking")
model3=c(model2,"creatinine","crp","hdlc","ldlc","triacylglycerols","sbp","dbp","hb")

```
### Unimputed models

```{r Regressions_Complete cases, results="asis"}
#models 1 and 2 for iron
iron.m1<-regression(model1,iron)
iron.m1 <- iron.m1 |> 
  mutate(model="Model 1")
iron.m2<-regression(model2,iron)
iron.m2 <- iron.m2 |> 
  mutate(model="Model 2")
#models 1 and 2 for ferritin
ferritin.m1<-regression(model1,ferritin)
ferritin.m2<-regression(model2,ferritin)
ferritin.m1 <- ferritin.m1 |> 
  mutate(model="Model 1")
ferritin.m2 <- ferritin.m2 |> 
  mutate(model="Model 2")

#models 1 and 2 for transferrin
transferrin.m1<-regression(model1,transferrin)
transferrin.m2<-regression(model2,transferrin)
transferrin.m1 <- transferrin.m1 |> 
  mutate(model="Model 1")
transferrin.m2 <- transferrin.m2 |> 
  mutate(model="Model 2")

#models 1 and 2 for tfsat
tfsat.m1<-regression(model1,tfsat)
tfsat.m2<-regression(model2,tfsat)
tfsat.m1 <- tfsat.m1 |> 
  mutate(model="Model 1")
tfsat.m2 <- tfsat.m2 |> 
  mutate(model="Model 2")

# combine models
complete_all <- bind_rows(iron.m1[2,1:12], iron.m2[8,1:12],
                          ferritin.m1[2,1:12], ferritin.m2[8,1:12],
                          transferrin.m1[2,1:12], transferrin.m2[8,1:12],
                          tfsat.m1[2,1:12], tfsat.m2[8,1:12])

#Save the regression models with unimputed data
saveRDS(complete_all,file=here::here("data_files","complete_all.RDS"))

#analysis for atherosclerotic plaque
unfilt_df <- readRDS(file=here::here("data_files","unfilt_df.RDS")) |> 
  mutate(plaque=if_else(imt>1.5,1,0),
         plaque=as.factor(plaque))

#filter for males
unfilt_m <- unfilt_df |> 
  filter(gender=="0") |>
  dplyr::mutate(iron = scale(iron), 
                ferritin = scale(ferritin), 
                transferrin = scale(transferrin), 
                tfsat = scale(tfsat))

#filter for females
unfilt_f <- unfilt_df |> 
  filter(gender=="1") |>
  dplyr::mutate(iron = scale(iron), 
                ferritin = scale(ferritin), 
                transferrin = scale(transferrin), 
                tfsat = scale(tfsat))

library(easystats)
plaque <- "plaque"
iron <- "iron"
age <- "age"
diabetes <- "diabetes"
ferritin <- "ferritin"
transferrin <- "transferrin"
tfsat <- "tfsat"
#males
iron_plaque.m <- glm(plaque~iron, family=binomial, data=unfilt_m) |>
  model_parameters(exponentiate = TRUE)

ferritin_plaque.m <- glm(plaque~ferritin, family=binomial, data=unfilt_m) |>
  model_parameters(exponentiate = TRUE)

transferrin_plaque.m <- glm(plaque~transferrin, family=binomial, data=unfilt_m) |>
  model_parameters(exponentiate = TRUE)

tfsat_plaque.m <- glm(plaque~tfsat, family=binomial, data=unfilt_m) |>
  model_parameters(exponentiate = TRUE)

#combine
plaque_M <- rbind(iron_plaque.m[2,1:9],ferritin_plaque.m[2,1:9],
                  transferrin_plaque.m[2,1:9],tfsat_plaque.m[2,1:9])

#females
iron_plaque.f <- glm(plaque~iron, family=binomial, data=unfilt_f) |>
  model_parameters(exponentiate = TRUE)

ferritin_plaque.f <- glm(plaque~ferritin, family=binomial, data=unfilt_f) |>
  model_parameters(exponentiate = TRUE)

transferrin_plaque.f <- glm(plaque~transferrin, family=binomial, data=unfilt_f) |>
  model_parameters(exponentiate = TRUE)

tfsat_plaque.f <- glm(plaque~tfsat, family=binomial, data=unfilt_f) |>
  model_parameters(exponentiate = TRUE)

#combine
plaque_F <- rbind(iron_plaque.f[2,1:9],ferritin_plaque.f[2,1:9],
                  transferrin_plaque.f[2,1:9],tfsat_plaque.f[2,1:9])

rm(iron_plaque.m,ferritin_plaque.m,transferrin_plaque.m,tfsat_plaque.m,
   iron_plaque.f,ferritin_plaque.f,transferrin_plaque.f,tfsat_plaque.f)

saveRDS(plaque_M,file=here::here("data_files","plaque_M.rds"))
saveRDS(plaque_F,file=here::here("data_files","plaque_F.rds"))


```


### Imputation
```{r Imputation, include=FALSE}

Y<-data_frame|>
  dplyr::select(!c(study, gender, hypertension, thalassemia, hemochromatosis, ckd,
              inflammation, anemia, age_gp,age_cat,hi_ferritin))
clus <- data_frame[, "study"]
nburn = as.integer(500)
nbetween = as.integer(500)
nimp = as.integer(5)
set.seed(58982)

m_imp <- jomo::jomo(Y = Y, clus = clus, nburn = nburn, nbetween = nbetween, nimp = nimp, meth = "common")

imp.mitml<-jomo2mitml.list(m_imp)


```


```{r Scale Imputed variables}
#scale variables for each variable in the list
imp.mitml[[1]] <- imp.mitml[[1]] |>
  mutate(iron = scale(iron), 
         ferritin = scale(ferritin), 
         transferrin = scale(transferrin), 
         tfsat = scale(tfsat), 
         age = scale(age), 
         creatinine = scale(creatinine), 
         hdlc = scale(hdlc), 
         ldlc = scale(ldlc), 
         triacylglycerols = scale(triacylglycerols),
         bmi = scale(bmi), 
         crp = scale(crp),
         sbp= scale(sbp),
         dbp=scale(dbp),
         hb=scale(hb))

imp.mitml[[2]] <- imp.mitml[[2]] |>
  mutate(iron = scale(iron), 
         ferritin = scale(ferritin), 
         transferrin = scale(transferrin), 
         tfsat = scale(tfsat), 
         age = scale(age), 
         creatinine = scale(creatinine), 
         hdlc = scale(hdlc), 
         ldlc = scale(ldlc), 
         triacylglycerols = scale(triacylglycerols),
         bmi = scale(bmi), 
         crp = scale(crp),
         sbp= scale(sbp),
         dbp=scale(dbp),
         hb=scale(hb))

imp.mitml[[3]] <- imp.mitml[[3]] |>
  mutate(iron = scale(iron), 
         ferritin = scale(ferritin), 
         transferrin = scale(transferrin), 
         tfsat = scale(tfsat), 
         age = scale(age), 
         creatinine = scale(creatinine), 
         hdlc = scale(hdlc), 
         ldlc = scale(ldlc), 
         triacylglycerols = scale(triacylglycerols),
         bmi = scale(bmi), 
         crp = scale(crp),
         sbp= scale(sbp),
         dbp=scale(dbp),
         hb=scale(hb))


imp.mitml[[4]] <- imp.mitml[[4]] |>
  mutate(iron = scale(iron), 
         ferritin = scale(ferritin), 
         transferrin = scale(transferrin), 
         tfsat = scale(tfsat), 
         age = scale(age), 
         creatinine = scale(creatinine), 
         hdlc = scale(hdlc), 
         ldlc = scale(ldlc), 
         triacylglycerols = scale(triacylglycerols),
         bmi = scale(bmi), 
         crp = scale(crp),
         sbp= scale(sbp),
         dbp=scale(dbp),
         hb=scale(hb))

imp.mitml[[5]] <- imp.mitml[[5]] |>
  mutate(iron = scale(iron), 
         ferritin = scale(ferritin), 
         transferrin = scale(transferrin), 
         tfsat = scale(tfsat), 
         age = scale(age), 
         creatinine = scale(creatinine), 
         hdlc = scale(hdlc), 
         ldlc = scale(ldlc), 
         triacylglycerols = scale(triacylglycerols),
         bmi = scale(bmi), 
         crp = scale(crp),
         sbp= scale(sbp),
         dbp=scale(dbp),
         hb=scale(hb))

saveRDS(imp.mitml,file = here::here("data_files","imputed_data.rds"))
beepr::beep()

```


\clearpage

## Session Information

```{r session-info, echo=FALSE} 
sessionInfo()
```
